% ============================================================
% Geographic Epistemic Asymmetries in LLM-Assisted Policy Research
% A 15-Country, 6-Model Benchmark
% ============================================================
% Working draft — pre-submission version
% Target journal: Patterns (Cell Press)
% Format: generic minimalist (adapted to Cell Press template at Stage 9)
% ============================================================

\documentclass[11pt,a4paper]{article}

% -- Essential packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

% -- Layout
\usepackage[margin=2.5cm]{geometry}
\usepackage{setspace}
\onehalfspacing
\usepackage{lineno}
\linenumbers

% -- Math and symbols
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}

% -- Graphics and tables
\usepackage{graphicx}
\graphicspath{{../figures/}}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{array}
\usepackage{caption}
\captionsetup{font=small,labelfont=bf}

% -- References
\usepackage[numbers,square,comma,sort&compress]{natbib}
\bibliographystyle{unsrtnat}

% -- Hyperlinks (load late)
\usepackage[colorlinks=true,
            linkcolor=blue!50!black,
            citecolor=blue!50!black,
            urlcolor=blue!50!black]{hyperref}

% -- Custom commands
\newcommand{\placeholder}[1]{\textcolor{red!80!black}{\textbf{[#1]}}}
\newcommand{\simvalue}[1]{\textcolor{blue!60!black}{\textit{#1}}}

% ============================================================
% METADATA
% ============================================================

\title{Geographic Epistemic Asymmetries in LLM-Assisted Policy Research:\\
A 15-Country, 6-Model, 4-Language Benchmark}

\author{
  Lucas Rover$^{1,*}$ \\
  Eduardo Tadeu Bacalhau$^{1}$ \\
  Yara de Souza Tadano$^{1}$ \\
  \\
  \small $^1$ Programa de P\'os-Gradua\c{c}\~ao em Planejamento Urbano e Ambiental,\\
  \small Universidade Tecnol\'ogica Federal do Paran\'a, Brazil\\
  \small $^*$ Corresponding author: \texttt{lucasrover@alunos.utfpr.edu.br}
}

\date{\today}

% ============================================================
% DOCUMENT
% ============================================================

\begin{document}

\maketitle

% Abstract — to be refined after Results are in
\input{sections/00_abstract}

\noindent\textbf{Keywords:} large language models; geographic bias; Global South;
epistemic justice; policy research; benchmark; multilingual NLP

\vspace{1em}

% Body
\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_methods}
\input{sections/04_results}
\input{sections/05_discussion}
\input{sections/06_conclusion}

% ============================================================
% BACK MATTER
% ============================================================

\section*{Data and code availability}

All code, prompts, ground-truth references, and (upon real execution) anonymised
LLM responses will be publicly available at
\href{https://github.com/Roverlucas/artigo-vies-analise-regional}{github.com/Roverlucas/artigo-vies-analise-regional}
under MIT (code) and CC-BY-4.0 (data/prompts/documentation) licenses. A permanent
Zenodo snapshot will be minted at submission. The analysis pipeline is
reproducible end-to-end via \texttt{python code/run\_all.py}.

\section*{Pre-registration}

The full statistical analysis plan was pre-registered at the Open Science
Framework at \placeholder{OSF DOI}. All confirmatory analyses reported here
follow the pre-registered specifications; exploratory analyses are labeled
explicitly and corrected separately.

\section*{Acknowledgements}

This work was supported by \placeholder{funding sources}. We acknowledge research
credit programs from OpenAI, Anthropic, and Google for partial support of API costs.

\section*{Author contributions}

Following CRediT taxonomy:
\begin{itemize}\setlength{\itemsep}{0pt}
\item \textbf{Lucas Rover:} Conceptualization, Methodology, Software, Data Curation,
  Formal Analysis, Investigation, Visualization, Writing~--~Original Draft,
  Project Administration.
\item \textbf{Prof.\ Dr.\ Eduardo Tadeu Bacalhau:} Validation, Supervision,
  Writing~--~Review \& Editing.
\item \textbf{Profa.\ Dra.\ Yara de Souza Tadano:} Validation, Supervision,
  Writing~--~Review \& Editing, Funding Acquisition.
\end{itemize}

Validation, supervision, and final manuscript review are shared by
Prof.\ Dr.\ Eduardo Tadeu Bacalhau and Profa.\ Dra.\ Yara de Souza Tadano.

\section*{Declaration of interests}

The authors declare no competing interests.

% Bibliography
\bibliography{references}

\end{document}
