\begin{abstract}

Large language models (LLMs) are increasingly used by researchers worldwide to
support applied work in public policy, yet the extent and mechanism of their
geographic factual bias in this specific use case remain unmeasured. We present
the first pre-registered benchmark designed to quantify LLM bias in
research-grade tasks about public policy, socioeconomic reality, and
environmental context across countries of the Global South. Our design combines
a theoretically-triangulated sample of 15 countries (stratified along UNCTAD
Global South/North, Joshi et al. (2020) linguistic-resource taxonomy, and World
Bank income groups), six frontier, open-source, and regionally-trained LLMs, a
sparse factorial matrix of prompt languages, and expert-authored prompts
validated by a multi-regional panel of domain specialists. We test four
pre-registered hypotheses: \textbf{(H1)} LLMs exhibit systematically lower
accuracy in the Global South (expected Cohen's \textit{d}~$\geq$~0.5);
\textbf{(H2)} prompt language and country interact non-additively;
\textbf{(H3)} a regional model (Sabi\'a-3) either closes the Global South gap
for its target country or displaces it onto linguistically-related countries;
\textbf{(H4)} corpus-representation proxies (Common Crawl tokens, Wikipedia
pageviews) mediate the geographic effect. \placeholder{Upon real execution:
insert main effect sizes with 95\% CIs}. \placeholder{Insert headline finding
of H3 — the most informative for the policy discussion.} Our findings
\placeholder{support / qualify / contradict} the coloniality-of-knowledge
hypothesis operationalized for AI systems, with direct implications for how
Global South researchers should use LLMs in their scholarly work and for policy
on sovereign AI investment. All prompts, code, ground-truth references, and
responses are openly available.

\end{abstract}
