\section{Introduction}
\label{sec:intro}

Large language models (LLMs) have moved from research curiosity to infrastructure
of scholarly work in under five years. Applied researchers — particularly those
working on public-policy questions — increasingly draft literature syntheses,
characterize institutional landscapes, identify stakeholders, and search for
primary sources with LLM assistance~\citep{messeri2024science,bail2024llm}. This
shift has happened faster than our ability to audit the tools it depends on.

Accumulating evidence shows that LLMs systematically under-represent knowledge
about the Global South. \citet{moayeri2024worldbench} documented 1.5$\times$
higher factual error rates for Sub-Saharan African countries versus North
American countries across 20 models and 11 World Bank indicators.
\citet{manvi2024llm} showed that LLM zero-shot ratings on subjective topics
(attractiveness, morality, intelligence) correlate strongly with income level
($\rho$ up to 0.70 against socioeconomic conditions).
\citet{mirza2024global} demonstrated that GPT-4 systematically privileges
statements about the Global North over the Global South in factual verification
tasks. These are important findings, but they share three limitations that
leave the central question for applied research unanswered.

\paragraph{Gap 1: benchmarks test the wrong tasks.} Existing evaluations use
numeric recall (WorldBench), subjective ratings (Manvi et al.), fact-checking
of journalistic claims (Global-Liar), or everyday cultural knowledge
(BLEnD~\citep{myung2024blend}). None tests the tasks that Global South
researchers actually delegate to LLMs: characterizing socio-environmental
context, synthesizing existing public-policy frameworks, identifying
institutional stakeholders, recommending primary sources. The practical
significance of geographic bias depends on its manifestation in these applied
settings, not in trivia.

\paragraph{Gap 2: language and geography are confounded.} BLEnD reports
non-monotonic effects of prompt language on accuracy, but its design --- 16
countries $\times$ 13 languages in everyday knowledge --- does not allow
isolation of the interaction. No study to date has performed a controlled
factorial test of language $\times$ country at the country level in a domain
relevant to applied research.

\paragraph{Gap 3: regional models are proposed as remedies, never tested as
such.} The release of regionally-trained LLMs (Sabi\'a-3 for Brazilian
Portuguese~\citep{almeida2025portuguese}; Latam-GPT for Spanish; AMALIA for
European Portuguese) has been accompanied by advocacy for ``sovereign AI''
investment. Whether these models close the Global-South performance gap or
merely displace it to neighbouring countries with shared linguistic heritage
(e.g., Angola, Mozambique for Portuguese; Peru's Indigenous populations for
Spanish) has never been empirically tested. The policy implications of these
two scenarios diverge sharply.

\paragraph{Gap 4: the mechanism is asserted, not tested.}
Underrepresentation in the training corpus is the dominant explanation for
geographic bias in the NLP literature~\citep{joshi2020state,bender2021stochastic}.
Yet no study has directly tested whether corpus-representation proxies
(Common Crawl token shares, Wikipedia pageviews by language) correlate
quantitatively with geographic accuracy gaps, while controlling for plausible
confounders such as GDP and digital infrastructure.

\subsection*{Contributions of this work}

We present the first pre-registered benchmark designed explicitly for these
four gaps. Our contributions are:

\begin{enumerate}\setlength{\itemsep}{2pt}
\item \textbf{Applied-research task suite.} We construct a suite of prompts
  authored by a multi-regional panel of domain experts (N~$\geq$~10 across
  Latin America, Africa, and Asia), covering three task types (direct factual
  recall, open characterization, source recommendation) across three research
  domains (public policy, socioeconomic reality, environmental context). Prompts
  simulate real tasks that Global South researchers delegate to LLMs.
\item \textbf{Theoretically-triangulated country sampling.} We select 15
  countries stratified along three independent axes: UNCTAD North/South
  classification (operationalising coloniality~\citep{mohamed2020decolonial}),
  Joshi et al.\ (2020) six-class linguistic-resource taxonomy, and World Bank
  income groups. This allows us to disentangle geographic, linguistic, and
  economic contributions to the bias for the first time.
\item \textbf{Head-to-head regional vs.\ frontier comparison.} We include
  Sabi\'a-3 (regionally trained on Brazilian Portuguese) alongside frontier
  proprietary (GPT-5, Claude Opus 4.7, Gemini 2.5 Pro) and open-source global
  (Llama~4, Qwen~3) models, testing pre-specified contrasts for whether
  regional training closes the gap or displaces it.
\item \textbf{Mechanistic test with sensitivity analysis.} We test the
  corpus-representation mediation hypothesis using log Common Crawl token
  shares and Wikipedia pageviews, with explicit DAG formalisation, E-value
  sensitivity analysis~\citep{vanderweele2017evalue}, and controls for GDP
  and digital infrastructure.
\item \textbf{Open resources.} All prompts (after panel validation), ground
  truth source documents, LLM responses, code, and figures are released under
  CC-BY-4.0 and MIT licenses, with a permanent Zenodo DOI and a companion
  Data Paper submitted to \textit{Scientific Data}.
\end{enumerate}

\subsection*{Theoretical framing}

We interpret our findings within the \emph{coloniality of knowledge}
framework~\citep{quijano2000coloniality,mohamed2020decolonial} and its recent
articulation for AI systems~\citep{couldry2019costs,kerche2026silicon}.
LLMs are not neutral tools; they are compressed statistical summaries of
corpora that were themselves produced under conditions of uneven digital
infrastructure, uneven publication access, and uneven political-economic
power. When those compressed summaries are then used by Global South
researchers to study their own societies, the resulting research risks
inheriting the biases of the corpus~\citep{dignazio2020data}. This positions
our empirical work as a test of one link in a theorised reproduction cycle
(training-corpus asymmetry $\to$ LLM knowledge asymmetry $\to$ research output
asymmetry), rather than as a narrow benchmark.

The paper proceeds as follows. Section~\ref{sec:related} positions our work
relative to the literature. Section~\ref{sec:methods} presents the
pre-registered design. Section~\ref{sec:results} reports the empirical
findings. Section~\ref{sec:discussion} interprets them through both statistical
and critical lenses and considers practical implications. Section~\ref{sec:conclusion}
closes.
