\section{Related Work}
\label{sec:related}

The literature on LLM bias consolidated in three overlapping waves that we
summarize below, identifying how our contribution differs from each.

\subsection{Wave 1 (2024): Discovery of factual geographic bias}

\citet{manvi2024llm} established that frontier LLMs carry systematic biases
against regions of lower socioeconomic conditions, demonstrating Spearman
correlations up to 0.70 between model-generated ratings of attractiveness,
morality, and intelligence and per-capita income. While compelling, the
outcome space (subjective ratings) leaves unclear whether the same bias
manifests in objective factual recall.

\citet{moayeri2024worldbench} introduced WorldBench, quantifying factual recall
against World Bank indicators across 20 LLMs and 11 indicators, documenting
1.5$\times$ higher absolute relative error for Sub-Saharan Africa versus North
America. WorldBench also pioneered automatic detection of citation hallucination
(models citing the World Bank while producing false statistics). Its design,
however, is bounded by the macro-economic indicator space; questions asked of
LLMs in applied policy research are typically more complex than ``what is the
GDP per capita of country $X$''.

\citet{mirza2024global} (Global-Liar) extended the question to temporal
stability, showing that newer GPT-4 versions do not monotonically improve and
that privileging of Global North statements is stable across versions. Their
contribution is temporal and across-version, not cross-model or cross-language.

Collectively, Wave 1 established that the phenomenon \emph{exists} and is
\emph{measurable}, but none of these studies addressed applied research use,
cross-model comparison including regional alternatives, or mechanistic
explanation.

\subsection{Wave 2 (2025): Multilingual and cultural extensions}

\citet{myung2024blend} (BLEnD) introduced 52k question-answer pairs covering 16
countries and 13 languages in everyday cultural knowledge. Their central
finding --- that LLMs perform better in native languages of high-resource
cultures but worse in native languages of low-resource ones --- motivates our
H2. Yet BLEnD's design conflates cultural-knowledge accuracy with research-use
accuracy, and its country selection is not stratified theoretically.

Regional benchmarks proliferated: BRoverbs~\citep{broverbs2025} for Brazilian
proverbs, TiEBe~\citep{almeida2025tiebe} for regional historical events,
ALBA~\citep{alba2026} and AMALIA~\citep{amalia2026} for European Portuguese.
These document concrete failures in culturally grounded tasks but do not
address applied-research usage, and each is limited to one or two languages.

\subsection{Wave 3 (2026): Alignment and mitigation}

\citet{opuszko2026unraveling} investigated whether reasoning models
(o1, DeepSeek-R1) reduce geographic bias, finding that reasoning helps in
aggregate but does not eliminate regional disparities. \citet{kerche2026silicon}
proposed a place-based typology of LLM biases, framing ours within a broader
sociotechnical landscape. These contributions are theoretical-conceptual; our
work provides empirical complementarity.

\subsection{LLMs in policy and research use}

A parallel literature examines LLM use in specific policy and research
contexts. \citet{he2025policymaking} evaluated LLM policy recommendations for
homelessness across four cities (Barcelona, Johannesburg, South Bend, and one
other), finding ``context-blind rigidity'' where LLMs apply stable internal
heuristics insufficiently calibrated to local sociopolitical realities. Their
work validates that the bias matters for policy; we generalize to a broader set
of countries, domains, and model categories with a pre-registered design.

In health specifically, \citet{kozlakidis2026medical} argue that LLM
hallucinations in clinical contexts pose acute challenges in regulatory
environments with limited technical capacity, highlighting that
``language and cultural mismatch between globally trained models and local
practice increases the risk of errors'' --- the mechanism our H2 tests
formally.

\subsection{Positioning of the present work}

Our design differs from prior work in four concrete ways: (i) applied-research
task design rather than numeric recall or cultural knowledge; (ii) country
selection stratified on three independent theoretical axes rather than
regional convenience; (iii) regional-model inclusion with pre-specified
contrasts rather than frontier-only comparison; (iv) explicit mediation test
of corpus-representation mechanism with sensitivity analysis. No prior study
combines these four elements.
