\section{Methods}
\label{sec:methods}

\subsection{Research design}

We conducted a pre-registered, factorial, multi-site benchmark experiment to
quantify geographic bias in six frontier and regional Large Language Models
across tasks representative of applied policy research in Global South
contexts. The design crossed 15 countries (stratified along three theoretically-
grounded axes), 6 LLMs, a sparse matrix of 4+ languages, 3 research domains,
and 5 task types, yielding approximately \simvalue{$\sim$23{,}400 unique
prompt-response pairs} after five stochastic replications per prompt-model pair.
The pre-registration was deposited at the Open Science Framework
(\placeholder{OSF DOI}) prior to any principal data collection and locked the
primary models, exclusion criteria, and multiple-comparison corrections.

\subsection{Country selection}

Countries were selected via theoretical
sampling~\citep{patton2015qualitative} triangulated across three independent
classifications. First, the \textbf{UNCTAD Global North/South
classification}~\citep{unctad2024classification} operationalizes the
coloniality-of-knowledge framework that motivates this
study~\citep{mohamed2020decolonial}. Second, the \textbf{Joshi et al. (2020)
six-class taxonomy} of linguistic-resource representation in NLP corpora
provided an orthogonal stratification dimension~\citep{joshi2020state}. Third,
\textbf{World Bank income groups (FY26)}~\citep{worldbank2025classification}
captured economic position independently of the previous two axes.

The final sample of 15 countries — Brazil, Mexico, Argentina, Peru (Latin
America); Nigeria, South Africa, Kenya, Egypt (Africa); India, Indonesia,
Bangladesh, the Philippines (Asia); and the United States, Germany, Japan
(Global North control) — covers four world regions, five Joshi classes
(1 through 5), and four World Bank income groups. Countries with compromised
statistical governance (e.g., active conflict) or without adequate public
ground-truth documentation were excluded; Oceania and Joshi Class~0 languages
were not represented given feasibility constraints and are acknowledged as
scope limitations.

\subsection{Model selection}

Six LLMs were benchmarked, representing three architectural categories.
\textbf{Three frontier proprietary models} --- GPT-5 (OpenAI), Claude Opus 4.7
(Anthropic), and Gemini 2.5 Pro (Google) --- were accessed via their official
APIs using the identical model version string throughout the collection window.
\textbf{Two open-weight globally-trained models} --- Llama 4 (70B) and Qwen 3
(32B) --- served as contrasts in architecture and training-data provenance,
with Qwen 3 functioning specifically as a scale-matched control for the
regional model. \textbf{One regionally-trained model} --- Sabi\'a-3 (Maritaca
AI), trained with substantial Brazilian Portuguese
emphasis~\citep{almeida2025portuguese} --- was included to test whether
regional training closes or reshapes the Global South performance gap.

\subsection{Prompt construction and validation}

A panel of \simvalue{2--3 domain experts per Global South region} was recruited
via the Graduate Program in Urban and Environmental Planning (PPGSAU/UTFPR), the
Latin American Council of Social Sciences (CLACSO), and the Council for the
Development of Social Science Research in Africa (CODESRIA), under a
collaboration protocol that included co-authorship and ethical approval
(Plataforma Brasil, \placeholder{CEP protocol number}). Experts constructed
prompts in three research domains --- (1) public policy and institutions,
(2) socioeconomic reality and demography, and (3) regional environmental
context --- across five task types: (T1) direct factual recall with verifiable
ground truth; (T2) descriptive characterisation through open-ended generation;
(T3) institutional stakeholder identification; (T4) primary-source
recommendation; and (T5) confidence calibration.

A pilot of 50 prompts was first validated for inter-rater agreement using
Krippendorff's $\alpha$, with a threshold of $\alpha \geq 0.70$ required
before proceeding to the final 450-prompt set. All prompts were authored in
English, then translated and back-translated by native-speaker professional
translators; any prompt whose back-translation produced semantic shift was
rewritten or excluded.

\subsection{Ground truth}

Ground truth for factual tasks was extracted from official government sources
(ministries, national statistics agencies, environmental agencies) and
peer-reviewed regional reports, with at least two independent sources per
prompt when available. A comprehensive catalog of 72+ sources across the 15
countries is provided as Supplementary Material. Extraction used custom
Python pipelines with OCR for scanned PDFs, and all sources were versioned by
access date, URL, and SHA-256 hash of the retrieved document to ensure
reproducibility. Where ground truth could plausibly have been seen by the
LLMs during training, a sensitivity subset was constructed using documents
dated after each model's training cutoff.

\subsection{Data collection}

API calls were executed at temperature~0.3 with deterministic seeds where
available, with five replications per prompt-model combination to estimate
stochastic variability. Collection occurred within a fixed four-week window to
minimize model-update confounds. All responses were stored alongside complete
request metadata (model version string, timestamp, token counts, finish
reason) in an immutable storage layer with SHA-256 checksums. Response quality
gates excluded truncated responses (\texttt{finish\_reason $\neq$ stop}) and
responses in languages divergent from the request; both were analyzed
separately as secondary outcomes.

\subsection{Measures}

The primary composite outcome variable combined five subcomponents with
pre-registered weights (see Supplementary Table~S1): (i) binary factual
accuracy against ground truth (for T1 tasks, weight 0.30); (ii) contextual
completeness scored against a pre-validated five-item rubric (for T2--T4
tasks, 0--5, weight 0.25); (iii) citation quality verified via DOI and source
check (0--5, weight 0.15); (iv) calibration error via Brier score (T5 tasks,
0--1, weight 0.15); and (v) absence of hallucination, coded binarily by human
raters (weight 0.15). Two independent raters, blinded to model identity,
evaluated each response using an Airtable-hosted rubric; inter-rater
disagreements were adjudicated by a third senior rater. Final Krippendorff's
$\alpha$ across raters reached \placeholder{observed value}.

\subsection{Corpus representation proxies}

For the mechanism test (H4), we extracted token counts per country from
Common Crawl (\placeholder{snapshot identifier locked in OSF}), aggregating
by country-code top-level domain as a lower-bound proxy, combined with
Wikipedia pageviews per language via the Wikimedia REST API. These proxies are
imperfect but constitute the standard measures used in prior
work~\citep{joshi2020state,almeida2025portuguese}; their limitations are
discussed in Section~\ref{sec:discussion}.

\subsection{Analytical strategy}

Primary analyses used Generalized Linear Mixed Models (GLMMs) with random
intercepts for country, model, and domain, implemented in Python via
\texttt{pymer4}~\citep{pymer4} which interfaces R's
\texttt{lme4}~\citep{bates2015lme4} through \texttt{rpy2}. For Gaussian
composite outcomes, \texttt{lmer} with REML estimation was used; for binary
outcomes (factual accuracy, hallucination), \texttt{glmer} with binomial
family and logit link was used. Hypothesis H1 was tested via the fixed effect
of country Global-South status on the composite outcome; H2 via the country
$\times$ language interaction term; H3 via pre-specified contrasts comparing
Sabi\'a-3 performance on Brazilian prompts against frontier-model average,
and Sabi\'a-3 performance on non-Brazilian Lusophone prompts to detect
displacement versus closure; H4 via country-level ecological regression of
composite accuracy on log(corpus tokens), followed by mediation analysis
using \texttt{semopy}~\citep{semopy} with bootstrap confidence intervals
(5{,}000 replicates).

All hypothesis tests used two-sided $\alpha = 0.05$ with Benjamini-Hochberg
FDR correction within each hypothesis family~\citep{benjamini1995fdr}. Monte
Carlo power analysis based on 2{,}000 simulated replicates per configuration
confirmed $>$95\% power for H1 at the design point (SESOI = 10 percentage
points, N = 15 countries), $\sim$85\% for H2 interaction,
and $\sim$77\% for H4 at $\rho = 0.60$ (the binding constraint; acknowledged
as a limitation).

Robustness analyses, all pre-registered, included: (i) leave-one-country-out
re-estimation; (ii) restriction to the expert-validated prompt subset;
(iii) Bayesian re-estimation with weakly informative priors via
\texttt{pymc}~\citep{pymc} to provide decision-theoretic posterior
probabilities; (iv) separate analysis excluding prompts with potential
training-data contamination. For the mediation analysis (H4), an E-value
sensitivity analysis~\citep{vanderweele2017evalue} quantified the strength of
unmeasured confounding required to nullify observed indirect effects.

\subsection{Pre-registration and open science}

The full protocol --- including prompts v1-final with hash, rubric, GLMM
specifications, multiple-comparison corrections, and exclusion criteria ---
was deposited at OSF prior to principal data collection. The analytic dataset,
after quality gates but before models were fit, was frozen and hashed; the
hash was committed to the OSF registry. All code, prompts, ground-truth
references, and anonymised response data are publicly available under
CC-BY-4.0 (data and prompts) and MIT License (code) at Zenodo with permanent
DOIs~\citep{zenodo2026dataset}, and a companion Data Paper is prepared for
\textit{Scientific Data}.

\subsection{Ethical considerations}

The expert-panel protocol received approval from UTFPR's Research Ethics
Committee via Plataforma Brasil (\placeholder{protocol}). Experts provided
informed consent and are acknowledged as co-authors when meeting ICMJE
criteria. No personal data from end-users of LLMs was collected. LLM API
terms of service were reviewed and respected; responses are redistributed
solely for research under fair-use and reproducibility provisions and
excluded from any commercial application.
