\section{Discussion}
\label{sec:discussion}

Our findings \placeholder{confirm / qualify / contradict} the hypothesis that
Large Language Models reproduce epistemic asymmetries along the Global
North--South divide in ways that directly affect their utility for applied
policy research. We discuss four interrelated implications: the empirical
contribution relative to prior work, the mechanistic evidence on corpus
representation, the policy implications of the regional-model finding, and
what the limitations tell us about where this literature should go next.

\subsection{What we learned empirically}

\subsubsection*{On H1 (geographic gap in applied research)}

\simvalue{The observed $\sim$10 percentage-point absolute gap} in accuracy
between Global South and Global North prompts on applied-research tasks is
broadly consistent in direction with WorldBench's 1.5$\times$ error ratio on
numeric indicators~\citep{moayeri2024worldbench} and Global-Liar's
privileging of Global North statements~\citep{mirza2024global}, but
importantly \emph{generalises} the finding to tasks that more closely
resemble real scholarly use. The pre-registered design, expert-authored
prompts, and panel-validated rubric distinguish this result from benchmarks
where task construction or evaluation is itself entangled with the bias
under study.

\subsubsection*{On H2 (language and geography interact)}

The non-monotonic interaction between prompt language and country, if
confirmed as hypothesised, extends the pattern first suggested in
BLEnD~\citep{myung2024blend} to applied-research tasks. For Global South
countries whose primary language is a \emph{high-resource} language in the
NLP sense (Brazilian Portuguese at Joshi Class~3; Spanish, Hindi at Class~4),
native-language prompts may provide a small advantage --- plausibly because
the LLM has absorbed more region-specific content in those languages. For
\emph{lower-resource} languages (Class~1--2), native-language prompts degrade
accuracy further, because the LLM lacks sufficient corpus to answer
competently and English remains its stronger knowledge channel.

This has an actionable implication: the often-repeated advice that
``LLMs work better in the local language'' is false for exactly the countries
that most need the tool --- those whose languages are under-represented in
the training corpus. Practitioners and researchers in countries of Joshi
classes 0--2 should default to English prompts; researchers in Joshi classes
3--4 countries may benefit from native-language prompts for regionally
specific questions.

\subsubsection*{On H3 (regional training)}

\simvalue{The observed advantage of Sabi\'a-3 on Brazilian prompts (+9pp
over frontier average, versus +3pp for scale-matched Qwen~3~32B)} is perhaps
the most policy-relevant finding. If confirmed empirically, it provides
evidence that regional training delivers a benefit \emph{beyond scale} ---
that is, a 65B-parameter model trained with Brazilian Portuguese emphasis
outperforms equally-sized globally-trained open-source models on the specific
country it targets.

The interpretation is nuanced, however. We do not know --- and cannot know
from this study --- whether the advantage extends to \emph{other Lusophone
countries} (Angola, Mozambique, Cabo Verde) or is specific to Brazil. If
specific to Brazil, ``regional'' training is effectively ``national''
training, and the policy conclusion is that each Global South country needs
its own nationally-tuned model. If it extends to Lusophone Africa, shared
linguistic-cultural heritage is a viable scaling axis. \placeholder{Our
preliminary Lusophone-Africa proxy test in Section~\ref{sec:results:h3}
suggests [result].}

\subsubsection*{On H4 (mechanism)}

The mediation analysis \simvalue{(62\% of the Global-South effect mediated
through corpus representation, E-value 2.3)} provides empirical content to
what has until now been a theoretical claim in the critical AI
literature~\citep{mohamed2020decolonial,couldry2019costs}. If corpus
representation is the primary mediator, then the Global-South accuracy gap
\emph{is largely remediable} --- through (a) deliberate expansion of Global
South content in training corpora, (b) regionally-tuned models as discussed
above, or (c) retrieval-augmented generation grounded in regional primary
sources. The E-value of 2.3 means that ruling out the mediation would
require an unmeasured confounder with associations of at least that strength
to both Global-South status and accuracy, which is plausible but not trivial
to construct.

\subsection{Theoretical implications}

Our results engage directly with the \emph{coloniality of knowledge}
literature. Quijano's~\citeyearpar{quijano2000coloniality} original argument
that global modernity was constituted through the systematic devaluation of
non-Western knowledge finds, in LLMs, a peculiarly \emph{technical}
manifestation: the compression of global text into a statistical model
preserves the pre-existing asymmetries of digital production. Mohamed et
al.'s~\citeyearpar{mohamed2020decolonial} call for decolonial AI predicted
this; our work quantifies it in a high-stakes applied-research domain.

Importantly, we do not conclude that LLMs are \emph{only} colonial
instruments. The regional-model finding suggests that deliberate technical
choices (which corpus to train on, in which languages, with what emphasis)
can meaningfully shift the distribution of bias. This is consistent with
D'Ignazio and Klein's~\citeyearpar{dignazio2020data} insistence that data
systems are always \emph{designed}: the question is whose knowledge is
designed in and whose is designed out.

\subsection{Practical implications}

For three audiences, the findings translate into specific guidance.

\paragraph{For Global South researchers using LLMs.}
\begin{itemize}\setlength{\itemsep}{0pt}
\item Do not use an LLM as the sole knowledge source for region-specific
  questions. The error rate is too high for unmediated use.
\item When a regional model exists and is well-calibrated for your country
  (e.g., Sabi\'a-3 for Brazil), prefer it for country-specific queries ---
  but do not assume it generalises across linguistic neighbours.
\item For countries whose primary language is Joshi Class 0--2, default to
  English prompts; for Joshi 3--5, native-language prompts may be acceptable
  or beneficial.
\item Treat LLM-provided citations, stakeholder lists, and source
  recommendations as \emph{hypotheses to verify}, not as reliable sources.
  Our hallucination rate for Global South citation queries was
  \placeholder{X\%}.
\end{itemize}

\paragraph{For model developers and research infrastructure.}
Current frontier proprietary models are better than global open-source
alternatives on absolute accuracy, but the gap versus Global North remains
substantial. Investment in Global South content for training corpora --- not
just in volume but in quality and in research-relevant documents
(policy texts, peer-reviewed regional scholarship, official statistics) ---
should be expected to reduce the measured bias, given our mediation result.

\paragraph{For public AI policy.}
The regional-model finding provides empirical support for policies that
invest in sovereign AI infrastructure, including Brazil's current interest
in Latam-GPT and analogous initiatives elsewhere. It also cautions against
over-generalising: a single regional model is unlikely to serve an entire
region uniformly. Policies should plan for \emph{portfolio} rather than
\emph{monolithic} investment.

\subsection{Limitations}

Five limitations deserve specific discussion.

\textbf{Synthetic versus real prompts.} Our prompts, while authored by
regional experts, remain synthetic: they were not drawn from live research
workflows. A logical next study is an auditing exercise on consented
research-use logs, complementing our controlled design.

\textbf{15 countries out of $\sim$195.} Our stratified sample covers four
continents and five Joshi classes but excludes Oceania, Joshi Class~0
languages, and countries with compromised statistical governance. Claims
generalize \emph{among stratification cells}, not universally.

\textbf{Corpus-representation mediation is correlational, not causal.} We
cannot manipulate corpus contents directly. Our E-value analysis quantifies
robustness but cannot rule out alternative mechanisms (tokenizer behaviour,
post-training alignment, political filtering) that could also contribute.

\textbf{Rubric is partially subjective.} Despite high inter-rater agreement,
the rubric for open-ended tasks (T2--T4) involves judgment calls. We mitigate
through panel validation and adjudication; we do not claim to have eliminated
evaluator subjectivity.

\textbf{Temporal snapshot.} LLMs update rapidly. Our results reflect a fixed
four-week collection window. The specific numerical values will shift; we
expect the qualitative pattern to persist until substantive interventions
on corpus composition are made.

\subsection{Directions for future work}

Our data leave three questions open that future studies should address:
(i) does the regional-model benefit extend to linguistic neighbours
(Lusophone Africa for Sabi\'a-3; Andean countries for a hypothetical
quechua-trained model)?; (ii) can retrieval-augmented generation grounded in
regional primary sources close the gap without full re-training?; and (iii)
does the bias manifest in downstream \emph{research outputs} when Global
South researchers use LLMs --- that is, does LLM-assisted writing in our
domain systematically introduce errors, citations, or framing biases that
map onto the patterns we measure here? Each is a natural next study.
