\section{Conclusion}
\label{sec:conclusion}

Large language models are becoming scholarly infrastructure. Their
geographic bias — now consistently documented across numeric recall, cultural
knowledge, and journalistic fact-checking — also manifests in exactly the
applied-research tasks that Global South scholars increasingly delegate to
them. Through a pre-registered benchmark covering 15 theoretically-stratified
countries, six LLMs spanning frontier proprietary, open-source global, and
regionally-trained categories, four prompt languages, three research domains,
and a multi-regional expert panel, we \placeholder{confirm/qualify/contradict}
the hypothesis that this bias is both (i) substantial in magnitude and (ii)
substantially mediated by training-corpus representation.

Three contributions distinguish this work. First, we extend the measurement
of LLM bias from trivia and cultural knowledge to the specific domain of
applied policy research, producing the first benchmark in which task design
reflects real scholarly use. Second, we demonstrate that regional training
\simvalue{(Sabi\'a-3)} delivers a benefit beyond scale for the country it
targets, while leaving open the empirical question of whether the benefit
generalises across linguistically-related countries. Third, we provide
empirical content to the coloniality-of-knowledge framework by quantifying
the proportion of geographic bias mediated through corpus representation
\simvalue{($\sim$62\%, E-value 2.3)}.

The practical implications are specific and actionable. For Global South
researchers: LLMs should not be sole sources for country-specific questions;
regional models should be preferred when available and well-calibrated for
the country in question; prompt language should be chosen based on the
target country's Joshi resource class. For developers and policymakers:
investment in corpus quality, not just volume, is a viable and empirically-
supported path to reducing bias; sovereign AI initiatives deserve empirical
attention rather than presumption of either their efficacy or their
limitations.

Most consequentially, our results point to a quiet form of epistemic
reproduction. LLMs trained on corpora produced under existing digital and
publication asymmetries, then used by Global South researchers to write
about their own societies, risk propagating those asymmetries into new
research. Interrupting that cycle requires not only better models but
deliberate choices about which knowledge the tools are trained to represent.
The data, code, and prompts accompanying this paper are released openly in
support of exactly those choices.
